---
title: ""
editor: visual
---

## Wisconsin

```{r,wisconsin,echo=FALSE, warning=FALSE,message=FALSE, fig.align='center'}
library(extrafont)
library(readxl)
library(ggplot2)
library(dplyr)
library(lubridate) 
WI_H2H <- read_excel("~/Documents/elections_files/swingstates.xlsx", 
    sheet = "WI_H2H")
#mean of the polls
wimean <- WI_H2H %>%
  dplyr::arrange(desc(party)) %>% 
  dplyr::group_by(answer) %>% 
  dplyr::mutate(pct_01da = zoo::rollmean(pct, k = 1, fill = NA),
                pct_02da =zoo::rollmean(pct, k =2, fill = NA),
                pct_03da = zoo::rollmean(pct, k = 3, fill = NA),
                pct_04da = zoo::rollmean(pct, k = 4, fill = NA),
                pct_06da = zoo::rollmean(pct, k = 6, fill = NA),
                pct_08da = zoo::rollmean(pct, k = 8, fill = NA),
                pct_10da = zoo::rollmean(pct, k = 10, fill = NA)) %>% 
  dplyr::ungroup()  
#changing the labels for the facets
wimean$TYPE <- factor(wimean$TYPE, levels = c("h_2_h", "multi"),
                  labels = c("Biden vs Trump", "Biden vs Trump with Third Parties")
                  )  
# plot 
ggplot(wimean, mapping = aes(x=end_date, y=pct_02da, color=answer))+
  geom_jitter() +
    geom_smooth(method = 'loess', formula = 'y~x',span=0.5, se=FALSE, lwd=1)+
  scale_colour_manual(values = c('#0099ff','#ff6666')) +
  labs(title = "President:General Election - Wisconsin 2024",color="",x="",y="",
       caption = "source::538.com\njamesstats.github.io/projectx")+
  theme_bw() +
  theme(legend.position = 'none',plot.caption = element_text(family = 'Andale Mono',face = 'italic'),
        strip.text.x = element_text(size = 12, color = "black", face = "bold.italic",family = "Andale Mono"),strip.text.y = element_text(size = 8,
        color = "black", face ="bold.italic",family = "Andale Mono"),strip.background = element_rect(color="black", fill="white", size=0.5, linetype="solid"
     ),plot.title = element_text(family = "Andale Mono",face = "bold"))+ 
  facet_wrap(~TYPE, nrow = 2) 
```

| End Date   | Biden | Trump     | Ballot                 |
|------------|-------|-----------|------------------------|
| May 9,2024 | 42.5% | 42.5%     | **Head-to-Head**       |
| May 9,2024 | 41.3% | **44.5%** | with **Third-Parties** |

*Used 2-Day weighted moving average(WMA) to aggregate the average of the polls.*

## Pennsylvania

```{r,pennsylvania,echo=FALSE, warning=FALSE,message=FALSE, fig.align='center'}

library(readxl)
library(ggplot2)
library(tidyverse) 
library(extrafont) 
library(dplyr)
library(lubridate) 
paH2H <- read_excel("~/Documents/elections_files/swingstates.xlsx", 
    sheet = "PA_H2H")
#sma
pamean <- paH2H %>%
  dplyr::arrange(desc(answer)) %>% 
  dplyr::group_by(answer) %>% 
  dplyr::mutate(pct_02da = zoo::rollmean(pct, k = 2, fill = NA),
                pct_03da = zoo::rollmean(pct, k = 3, fill = NA),
                pct_04da = zoo::rollmean(pct, k = 4, fill = NA),
                pct_06da = zoo::rollmean(pct, k = 6, fill = NA),
                pct_07da = zoo::rollmean(pct, k = 7, fill = NA),
                pct_08da = zoo::rollmean(pct, k = 8, fill = NA),
                pct_10da = zoo::rollmean(pct, k = 10, fill = NA)) %>% 
  dplyr::ungroup() 
#changing the labels for the facets
pamean$TYPE <- factor(pamean$TYPE, levels = c("h_2_h", "multi"),
                  labels = c("Biden vs Trump", "Biden vs Trump with Third Parties")
                  )  
# plot 
ggplot(pamean, mapping = aes(x=end_date, y=pct_02da, color=answer))+
  geom_jitter()+
    geom_smooth(method = 'loess', formula = 'y~x',span=0.5, se=FALSE, lwd=1)+
  scale_colour_manual(values = c('#0099ff','#ff6666')) +
  scale_y_continuous(limits = c(min(30),max(50)))+
  labs(color="",x="",y="",title = "President:General Election - Pennsylvania 2024",
       caption = "source::538.com\njamesstats.github.io/projectx")+
  theme_bw()+
  theme(legend.position = 'none',plot.caption = element_text(family = 'Andale Mono',face = 'italic'),
        strip.text.x = element_text(size = 12, color = "black", face = "bold.italic",family = "Andale Mono"),strip.text.y = element_text(size = 8,
        color = "black", face ="bold.italic",family = "Andale Mono"),strip.background = element_rect(color="black", fill="white", size=0.5, linetype="solid"
     ),plot.title = element_text(family = "Andale Mono",face = "bold"))+ 
  facet_wrap(~TYPE, nrow = 2)  

  
```

| End Date      | Biden | Trump     | Ballot                 |
|---------------|-------|-----------|------------------------|
| April 29,2024 | 40.5% | **44%**   | **Head-to-Head**       |
| April 29,2024 | 42.8% | **45.8%** | with **Third-Parties** |

*Used 2-Day weighted moving average(WMA) to aggregate the average of the polls.*

## Michigan

```{r,michigan,echo=FALSE, warning=FALSE,message=FALSE, fig.align='center'}

library(readxl)
library(extrafont) 
library(ggplot2)
library(dplyr)
library(lubridate) 
#fetching the data
miH2H <- read_excel("~/Documents/elections_files/swingstates.xlsx", 
    sheet = "MI_H2H")
#sma
mimean <- miH2H %>%
  dplyr::arrange(desc(answer)) %>% 
  dplyr::group_by(answer) %>% 
  dplyr::mutate(pct_02da = zoo::rollmean(pct, k = 2, fill = NA),
                pct_03da = zoo::rollmean(pct, k = 3, fill = NA),
                pct_04da = zoo::rollmean(pct, k = 4, fill = NA),
                pct_06da = zoo::rollmean(pct, k = 6, fill = NA),
                pct_08da = zoo::rollmean(pct, k = 8, fill = NA),
                pct_10da = zoo::rollmean(pct, k = 10, fill = NA),
                pct_10da = zoo::rollmean(pct, k = 12, fill = NA)) %>% 
  dplyr::ungroup() 
#renaming the labels
mimean$TYPE <- factor(mimean$TYPE, levels = c("h_2_h", "multi"),
                  labels = c("Biden vs Trump", "Biden vs Trump with Third Parties")
                  )  
# plot 
ggplot(mimean, mapping = aes(x=end_date, y=pct_02da, color=answer))+
  geom_jitter() +
    geom_smooth(method = 'loess', formula = 'y~x',span=0.5, se=FALSE, lwd=1)+
  scale_colour_manual(values = c('#0099ff','#ff6666')) +
  scale_y_continuous(limits = c(min(30),max(50)))+
  labs(title = "President:General Election - Michigan 2024",color="",x="",y="",
       caption = "source::538.com\njamesstats.github.io/projectx")+
  theme_bw() +
  theme(legend.position = 'none',plot.title = element_text(family = 'Andale Mono',face = 'bold'),
        plot.caption = element_text(family = 'Andale Mono',face = 'italic'),
        strip.text.x = element_text(size = 12, color = "black", face = "bold.italic",family = "Andale Mono"),strip.text.y = element_text(size = 8,
        color = "black", face ="bold.italic",family = "Andale Mono"),strip.background = element_rect(color="black", fill="white", size=0.5, linetype="solid"))+ 
  facet_wrap(~TYPE, nrow = 2) 
```

| End Date      | Biden | Trump     | Ballot                 |
|---------------|-------|-----------|------------------------|
| April 29,2024 | 41.8% | **43%**   | **Head-to-Head**       |
| April 29,2024 | 43.5% | **45.3%** | with **Third-Parties** |

*Used 2-Day weighted moving average(WMA) to aggregate the average of the polls.*
